# File: README.Rmd
---
output: github_document
---

# modelscompete4

An R package for comparing latent variable models with enhanced parameter extraction and model fit assessment.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# Install from GitHub (if publicly available)
# devtools::install_github("yourusername/modelscompete4")
```

## Features

Compare multiple SEM models simultaneously

Support for both nested and non-nested models

Chi-square difference tests for nested model comparisons

Extract multiple fit indices: AIC, BIC, CFI, TLI, RMSEA, SRMR

User-friendly interface with automatic model naming

Comprehensive results with structured output

Statistical tests: LRT (Likelihood Ratio Test) for nested models

## Quick Start

``` r
library(lavaan)
library(modelscompete4)

# Load example data
data(HolzingerSwineford1939)
HS_data <- HolzingerSwineford1939[, paste0("x", 1:6)]

# Define models
model1 <- 'Visual =~ x1 + x2 + x3
           Textual =~ x4 + x5 + x6
           Visual ~~ Textual'

model2 <- 'Visual =~ x1 + x2 + x3
           Textual =~ x4 + x5 + x6
           Visual ~~ 0*Textual'

# Fit models
fit1 <- cfa(model1, data = HS_data)
fit2 <- cfa(model2, data = HS_data)

# Compare models
results <- compare_models(
  fit1, fit2,
  model_names = c("Correlated", "Uncorrelated"),
  type = "nested",
  test = "lrt"
)

print(results)
```

## Example

### Non-Nested Model Comparison

``` r
# Compare three different factor structures
model_a <- 'F1 =~ x1 + x2 + x3'
model_b <- 'F1 =~ x1 + x2 + x3 + x4'
model_c <- 'F1 =~ x1 + x2 + x3 + x4 + x5'

fit_a <- cfa(model_a, data = HS_data)
fit_b <- cfa(model_b, data = HS_data)
fit_c <- cfa(model_c, data = HS_data)

results <- compare_models(
  fit_a, fit_b, fit_c,
  model_names = c("3 Indicators", "4 Indicators", "5 Indicators"),
  type = "non_nested",
  criteria = c("AIC", "BIC", "CFI", "RMSEA")
)
```

### Advanced Comparison with Multiple Fit Indices

``` r
results_advanced <- compare_models(
  fit1, fit2, fit3,
  model_names = c("Model1", "Model2", "Model3"),
  type = "non_nested",
  criteria = c("AIC", "BIC", "CFI", "TLI", "RMSEA", "SRMR"),
  test = "none",
  alpha = 0.05
)
```
## Function Reference

### compare_models()
The main function for model comparison:

``` r
compare_models(
  ...,                    # lavaan model objects
  model_names = NULL,     # Optional names for models
  type = "non_nested",    # "nested" or "non_nested"
  criteria = c("AIC", "BIC"),  # Fit indices to extract
  test = "none",          # "none", "lrt", or "wald"
  alpha = 0.05            # Significance level
)
```

### Returns: An object of class modelscompete4_advanced containing:

fit_indices: Data frame of fit indices for all models
diff_tests: Chi-square difference tests (for nested models)
models: Original lavaan model objects
comparisons: Detailed comparison data

## Output Structure

``` r
# Access different components of results
results$fit_indices      # Fit indices data frame
results$diff_tests       # Chi-square difference tests
results$models           # Original lavaan models
results$comparisons      # Detailed comparison data
```

### Vignette

For detailed examples and tutorials, see the package vignette:

``` r
vignette("modelscompete4-tutorial", package = "modelscompete4")
```

### Contributing

This package is licensed under the GNU General Public License v3.0 (GPL-3).


### License

This package is licensed under the GNU General Public License v3.0 (GPL-3).

## modelscompete4 Tutorial

### Introduction

The `modelscompete4` package provides tools for comparing multiple Structural Equation Models (SEM). This tutorial will guide you through the main features.

### Installation

``` r
# Install from GitHub
# devtools::install_github("ssjerf-stack/modelscompete4")

# Load the package
library(modelscompete4)
```
## Basic Usage

### Example 1: Simple Model Comparison

```{r, basic_usage}
# Load example data
data(HolzingerSwineford1939)
HS_data <- HolzingerSwineford1939[, paste0("x", 1:6)]

# Define two models
model1 <- 'Visual =~ x1 + x2 + x3
           Textual =~ x4 + x5 + x6'
           
model2 <- 'Factor =~ x1 + x2 + x3 + x4 + x5 + x6'

# Fit models
fit1 <- cfa(model1, data = HS_data)
fit2 <- cfa(model2, data = HS_data)

# Compare models
results <- compare_models(fit1, fit2)
print(results)
```

### Example 2: Nested Model Comparison with Chi-Square Test

```{r, nested_model_comparison}
# Nested models (correlated vs uncorrelated factors)
model_correlated <- '
  Visual =~ x1 + x2 + x3
  Textual =~ x4 + x5 + x6
  Visual ~~ Textual
'

model_uncorrelated <- '
  Visual =~ x1 + x2 + x3
  Textual =~ x4 + x5 + x6
  Visual ~~ 0*Textual
'

fit_correlated <- cfa(model_correlated, data = HS_data)
fit_uncorrelated <- cfa(model_uncorrelated, data = HS_data)

# Compare with chi-square test
results_nested <- compare_models(
  fit_correlated, fit_uncorrelated,
  model_names = c("Correlated", "Uncorrelated"),
  type = "nested",
  test = "lrt"
)

print(results_nested)
```

### Advanced Features
Custom Fit Indices
You can specify which fit indices to extract:

```{r, fit_indices}
# Compare with multiple fit indices
results_advanced <- compare_models(
  fit1, fit2,
  model_names = c("Two-Factor", "One-Factor"),
  criteria = c("AIC", "BIC", "CFI", "TLI", "RMSEA", "SRMR"),
  test = "none"
)

print(results_advanced)
```

### Comparing Multiple Models

```{r, compare_multiple_models}
# Create three different models
model_a <- 'F1 =~ x1 + x2 + x3'
model_b <- 'F1 =~ x1 + x2 + x3 + x4'
model_c <- 'F1 =~ x1 + x2 + x3 + x4 + x5'

fit_a <- cfa(model_a, data = HS_data)
fit_b <- cfa(model_b, data = HS_data)
fit_c <- cfa(model_c, data = HS_data)

# Compare all three
results_multi <- compare_models(
  fit_a, fit_b, fit_c,
  model_names = c("3 Indicators", "4 Indicators", "5 Indicators"),
  type = "non_nested"
)

print(results_multi)
```

### Working with Results

##### Accessing Components

```{r, accessing_components}
# Get fit indices table
fit_table <- results_multi$fit_indices
print(fit_table)

# Get model objects
models_list <- results_multi$models

# Get detailed comparisons
detailed <- results_multi$comparisons
```

### Exporting Results

```{r, accessing_components}
# Save results to file
saveRDS(results_multi, "model_comparison_results.rds")

# Export fit indices to CSV
write.csv(results_multi$fit_indices, "fit_indices.csv", row.names = FALSE)
```

### Best Practices

- Always check model convergence before comparison
- Use meaningful model names for clear output
- Consider both statistical significance and practical fit when interpreting results
- For nested models, use chi-square difference tests
- For non-nested models, rely on information criteria (AIC, BIC)

### Troubleshooting

#### Common Issues

1. Models not converging: Check model specification and data
2. Warning messages: Review model fit and identification
3. Chi-square test failing: Ensure models are properly nested

## Conclusion

The `modelscompete4` package provides a comprehensive toolkit for SEM model comparison. With support for both nested and non-nested models, multiple fit indices, and statistical tests, it streamlines the model selection process.

For more information, see the package documentation:

```r
?compare_models
help(package = "modelscompete4")
```
